### 2014-06-06 paper/abstract.tex
added todo command

% !TEX root = main.tex

\edit{write this abstract}