### 2014-06-06 paper/abstract.tex
edits from jason and yoshua

% !TEX root = main.tex

\edit{write this abstract}

...

We make the first comparison because \cite{Jarrett-ICCV2009} showed, --- quite strikingly --- that the combination of random convolutional features,  rectification, pooling, and local normalization can work almost as well as learned features. They showed this on relatively small networks

Our results from layers 1 and 2 suggest that getting random features to work well may not be as straightforward as they found. We use max pooling and local normalization on layers 1 and 2 just as they did, but we use relu(x) where they used abs(tanh(x)). The difference in performance could be due to a few factors: slightly different nonlinearity, the dataset itself, sizes of critical pieces of the network, etc.